---
title: Generalized Additive Model
subtitle: study group
author: "Hsiao-Hang & Oscar"
institute: "Institute of Oceanography, National Taiwan University"
date: "12/17/2019"

#titlegraphic: /Dropbox/teaching/clemson-academic.png
fontsize: 14 pt
output:
 beamer_presentation:
    theme: "AnnArbor"
    #template: svm-latex-beamer.tex
    keep_tex: true
    latex_engine: xelatex # pdflatex also works here
    dev: cairo_pdf # I typically comment this out  if latex_engine: pdflatex
    slide_level: 2
#make149: true
#mainfont: "Open Sans" # Try out some font options if xelatex
#titlefont: "Titillium Web" # Try out some font options if xelatex

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

